#!/usr/bin/env Rscript

library(rmarkdown)
render("hw.rmd", "pdf_document")

<!---
#library(knitr)
# TODO: YAML metadata removes the need for format
#pandoc('hw.rmd', format='latex', clean=TRUE, encoding=utf8, run_pandoc=TRUE)
# run_pandoc()

#     output:
#       pdf_document:
#         toc: true
#         highlight: zenburn
#       html_document:
#         toc: true
#         theme: united

# TODO: germaneca


# TODO: how to NOT print the above commands to the resulting .pdf?
# TODO: how to avoid putting everyting from here on in ""s?
# TODO: how to avoid mentioning the file name above?
# TODO: how to render special symbols, such as tilde, miu, sigma?
#     Unicode character (U+3BC) not set up for use with LaTeX.
#     See the inputenc package documentation for explanation.
-->

[comment]: <> (nano hw.rmd  && ./hw.rmd && evince hw.pdf)

"
4E1. In the model definition below, which line is the likelihood?
A: y_i is the likelihood, based on the expectation and deviation.


4E2. In the model definition just above, how many parameters are in the posterior distribution?
A: Two parameters, zero predictor variables.


4E3. Using the model definition above, write down the appropriate form of Bayes’ theorem that
includes the proper likelihood and priors.
A: P(miu, sigma | data) = P(data | miu, sigma) * P(p) / P(miu, sigma)
Numerator is summed over data point indices.
Denominator is summed once for every of the two parameters in the joint distribution.


4E4. In the model definition below, which line is the linear model?
A: miu_i = alpha + beta_i ... an affine model really


4E5. In the model definition just above, how many parameters are in the posterior distribution?
A: We have one predictor variable.
We have 3 parameters, because this is the count of numbers we have to teach the model.


4M1. For the model definition below, simulate observed heights from the prior (not the posterior).
A:
```{r}
points <- 10
rnorm(points, mean=rnorm(points, 0, 10), sd=runif(points, 0, 10))
```
4M3. Translate the map model formula below into a mathematical model definition.
A:
```{r}
flist <- alist(
y ~ dnorm( mu , sigma ),
miu ~ dnorm( 0 , 10 ),
sigma ~ dunif( 0 , 10 )
)
```
"
