#!/usr/bin/env Rscript


library(rmarkdown)
render("hw.rmd", "pdf_document", run_pandoc=TRUE, encoding="UTF-8")  #  All those parameters are not needed.
stopifnot(FALSE)  # Stop trying to interpret the rest of the file as valid R.
                  # Using this instead of `stop()` for the more obvious error message.

---
title: "Multivariate Linear Models"
author: Miroslav Vitkov
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: true
---


<!---
TODO: how to NOT print the above commands to the resulting .pdf?
TODO: how to avoid putting everyting from here on in ""s?
TODO: how to avoid mentioning the file name above?
TODO: how to render special symbols, such as tilde, miu, sigma?
    Unicode character (U+3BC) not set up for use with LaTeX.
    See the inputenc package documentation for explanation.


nano hw.rmd  && ./hw.rmd && evince hw.pdf)


"
5E1. Which of the linear models below are multiple linear regressions?
(1) μ_i = α + β_x,i
(2) μ_i = β_x x_i + β_z z_i
(3) μ_i = α + β(x_i − z_i)
(4) μ_i = α + β_x x_i + β_z z_i
"
-->

"  
5E1  
(1) - no, only a single predictor  
(2) - yes, intercept == 0  
(3) - yes, bx == bz  
(4) - yes  
"

"

# Style improved compared to last week:  

4M1. For the model definition below, simulate observed heights from the prior (not the posterior).
A:
```{r}
points <- 10
rnorm(points, mean=rnorm(points, 0, 10), sd=runif(points, 0, 10))
```
4M3. Translate the map model formula below into a mathematical model definition.
A:
```{r}
flist <- alist(
y ~ dnorm( mu , sigma ),
miu ~ dnorm( 0 , 10 ),
sigma ~ dunif( 0 , 10 )
)
```
"
