#!/usr/bin/env Rscript


#<!---
library(rmarkdown)
render("hw.rmd", output_format="pdf_document")
stopifnot(FALSE)  # Stop trying to interpret the rest of the file as valid R.
#-->


---
title: "Multivariate Linear Models"
author: Miroslav Vitkov
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: true
---


<!---
TODO: how to avoid putting everyting from here on in ""s?
TODO: how to avoid mentioning the file name above?
TODO: output format "all" doesn't read the YAML header

nano hw.rmd  && ./hw.rmd && evince hw.pdf)
-->


"  
5E1.  
Which of the linear models below are multiple linear regressions?  
(1) $\mu[i] = \alpha + \beta x[i]$  
(2) $\mu[i] = \beta[x] x[i] + \beta[z] z[i]$  
(3) $\mu[i] = \alpha + \beta (x[i] - z[i]$  
(4) $\mu[i] = \alpha + \beta[x] x[i] + \beta[z] z[i]$  
A:  
(1) - no, only a single predictor  
(2) - yes, intercept == 0  
(3) - yes, bx == bz  
(4) - yes  

# Style improved compared to last week:  

4M1. For the model definition below, simulate observed heights from the prior (not the posterior).
A:
```{r}
points <- 10
rnorm(points, mean=rnorm(points, 0, 10), sd=runif(points, 0, 10))
```
4M3. Translate the map model formula below into a mathematical model definition.
A:
```{r}
flist <- alist(
y ~ dnorm( mu , sigma ),
miu ~ dnorm( 0 , 10 ),
sigma ~ dunif( 0 , 10 )
)
```
"
