\documentclass{article}
\title{Project Proposal: Face recognition in real time; learning new faces in real time}
\author{'Miroslav Vitkov'}
\begin{document}
\maketitle

\section{Goal}
This is an engineering/scientific project with privacy and security implications.

\subsection{Primary Goals}
One primary goal is to manufacture an autonomous computational system.
It is to consist of a generic software component and a vaguely specified hardware component.
It must be robust(not crash) and exhibit false positive rate below 1 in 20 encounters.
\par
Another primary goal is to explore which face recognition model or implementation is best suited to continuous updating.
Scalability is paramount, but the format of this project doesn't provide fo such an advanced concept.

\subsection{Secondary targets}
\begin{itemize}
    \item{Achieve reasonable performance in low-light conditions - a true positive at least 9 times out of 10 encounters.}
    \item{Demonstrate system stability of at least MTBF=90 days.}
    \item{Be cross-platform and free as in beer and as in speech.}
    \item{Employ a minimal amount of reasonably popular tools to build and run.}
    \item{Run on an old i3 processor, no dedicated graphics and a 720p camera}
\end{itemize}

\section{Hypothesis}
I expect the current state of art, Haar features, to be most practical.
This will be tested by fitting it and at least 2 more models to the training data.
Emphasis must be placed both on model selection and model training.
\par
Evaluation is going to be performed by comparing precision and recall scores over a single encounter with the same scores over a single image in literature.
This is an optimistic approach but is somewhat counteracted by the bad operating conditions of the system.


\section{Training Data}
The training data will be collected privately and NOT published.
The product's performance can subsequently be tested on proliferated datasts, on which it is going to provide pessimistic performance values.

\section{Encounter}
An encounter is defined as 2 continuous seconds during which the subject's face is exposed to the camera.

\section{Conclusion}
This project seeks to provide two deliverable.
One is a functional software system.
The other is an objective comparison of face recognition algorithms in the context of learning new faces real-time.
\end{document}
